File-based localization of numerical perturbations in data analysis pipelines

### 2020-03-31 paper.tex
Edits in method

\toolname identifies the processes in a pipeline that produce different results when executed in two different conditions.  
In Condition 1, a directed, rooted, bipartite provenance graph is recorded for
each subject, where nodes represent application processes and files, and
edges represent read and write file accesses (Step 1 in Figure~\ref{fig:overview-tool}).
\emph{Transient} files are then identified as
 files that are either deleted during the execution, or modified by
multiple processes (Step 2). \toolname re-runs the pipeline to save these files
right after they are produced, resulting in a provenance \emph{DAG}
(Directed Acyclic Graph) in which file nodes have a single parent (in-degree
of 1) (Step 3) \tristan{Ali, this is not true in Figure 1: at step 3 there are still transient files}. This DAG is then compared to the one produced from
Condition 2, in a step-by-step execution that prevents the propagation of differences (Step 4).
\toolname finally produces a labeled graph that identifies the
non-reproducible processes of the pipeline.

\toolname makes the following assumptions on the pipeline, meant to ensure that 
a file can be unambiguously associated with the process that created it:
\begin{enumerate}
  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported.
  \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
  \item Temporary files used by multiple processes are not removed by the process that created them. \tristan{explain}
\end{enumerate} 
In addition, \toolname also requires that processes are associated to a
command line (executable and arguments), to facilitate process instrumentation.

@@

\tristan{The text switches between active and passive voice, this should be fixed.}

@@

\toolname uses the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}-left).

Provenance graphs are often subject-dependent.
%, which justifies the processing
% of multiple subjects. For instance, data acquisitions may be
% repeated a different number of times in each subject, for noise-reduction
% purposes. 
Some of these differences can be neglected, for instance when a
data decompression step is present at the beginning of the execution for
some subjects only, but other differences cannot, for instance
when complete different processing paths are used in different subjects.

@@

\begin{figure}
\centering
  \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
  \caption{Provenance graphs
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes are represented with circles and files with rectangles.
  Transient files are shaded in gray. 
  Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls) 
  \tristan{Does spot make any use of the process tree at any stage? If not, 
  I think we may want to remove these edges. 
  Ali: Yes, it uses process tree structure to do some tree operations like search, filter, or remove.
  Also using the process tree we do clustering later}. 
  Dashed edges represent file dependencies.
  Every node in the graph is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process or the file name.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  the rest of processes are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 73 is forked by process 2: 
  it was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}.  
}
  \label{fig:simple_script}
\end{figure}

@@

\toolname captures temporary files by replacing every
process P by a wrapper that first calls P and then saves the produced
temporary files to a read-only directory. This process replacement does not require 
modifying the pipeline: it only pre-pends to the \texttt{PATH} environment
variable a directory containing a wrapper script named after the executable
called by P.

Files written by multiple processes are disambiguated as follows. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, it
 replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way,
 successive versions of F are preserved for comparison. \toolname finally
 updates the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1 (Figure~\ref{fig:simple_script}-right). This operation also makes the provenance graph
 acyclic, since we assumed that a process could only release a single version of a file.

As an example, in the sample pipeline in
Listing~\ref{listing:sample-script}, file \emph{output.nii.gz} is first written
by \texttt{73\#bet2} and then overwritten by process \texttt{75\#fslmaths} as a temporary file. 
\tristan{add this comment to caption}

@@

Once transient files are captured, \toolname labels pipeline processes
through a step-by-step execution in Condition 2. Each process in Condition
2 is run and its output files are compared to the ones produced in
Condition 1. If no differences are found, the process is marked as
reproducible. Otherwise, the process is marked as non-reproducible and the
output files produced in Condition 1 are copied to Condition 2, to ensure
that differences do not propagate further in the pipeline. Processes are
instrumented transparently, through a modification of the \texttt{PATH}
variable similar to the one described previously. By default, differences
in output files are identified by comparing file checksums. Other
comparison functions can also be defined for specific file types, for
instance to ignore sections of the file containing timestamps or other
execution-specific information. Finally, \toolname creates a labeled
provenance graph, highlighting non-reproducible processes.

@@

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. 
At step one, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{73\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
Condition 1.
At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
as reproducible (green) as it produced files without differences.
At step 3, \texttt{76\#fslstats} is also labeled as 
reproducible as the last process of the pipeline.

@@

>>>>

\toolname identifies the processes in a pipeline that produce different
results when executed in two different conditions. First, a directed
bipartite provenance graph is recorded for each subject, where nodes
represent application processes and files, and edges represent read and
write file accesses (Step 1 in Figure~\ref{fig:overview-tool}). Second,
transient files, i.e., files that are either deleted during pipeline
execution or modified by multiple processes, are identified (Step 2) and
disambiguated, resulting in a provenance \emph{DAG} (Directed Acyclic
Graph) in which file nodes have a single parent (in-degree of 1) (Step 3)
\tristan{Ali, this is not true in Figure 1: at step 3 there are still
transient files}. DAGs produced in the two conditions are then compared, in
a step-by-step execution that prevents the propagation of differences in the pipeline (Step
4). Finally, a labeled graph is produced that identifies the
non-reproducible processes in the pipeline.

To ensure that a file can be unambiguously associated with the process that
created it, we assume that pipeline processes can be re-ordered such that:
\begin{enumerate}
\item Processes don't run concurrently;
\item Each process sequentially reads, computes, and writes. 
\end{enumerate}
In practice, pipeline processes may still run concurrently provided that
they don't write concurrently to the same files. A process may also
interleave file writes with computing, for instance when different file
blocks are processed sequentially. However, only a single version of the
file must eventually be made available to the other processes. Similarly,
in case processes write and delete temporary files throughout their
execution, these temporary files must not be read by any other process in
the pipeline.

In addition, we also require that processes are
associated to a command line (executable and arguments), to facilitate
process instrumentation.

@@

@@

We use the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
Using the \reprozip trace, \toolname reconstructs a provenance graph by creating process and file
nodes and by adding directed edges corresponding
to file reads and writes (Figure~\ref{fig:provenance-graph}).

Provenance graphs are often subject-dependent, due to variations in input data among subjects.
%, which justifies the processing
% of multiple subjects. For instance, data acquisitions may be
% repeated a different number of times in each subject, for noise-reduction
% purposes. 
Some of these differences can be neglected, for instance when a
data decompression step is present at the beginning of the execution for
some subjects only, but other differences cannot, for instance
when complete different processing paths are used in different subjects. \tristan{We need to conclude on that comment}

@@

\begin{figure}
\begin{subfigure}{0.4\linewidth}
  [ FIGURE GOES HERE ]

  \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
  \label{fig:provenance-graph}
\end{subfigure}
\hfill
\begin{subfigure}{0.4\linewidth}
  [ FIGURE GOES HERE ]

  \caption{Provenance DAG, with disambiguated transient files.}
  \label{fig:provenance-dag}
\end{subfigure}
  % \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
  \caption{Provenance graphs
  created from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes are represented with circles, files with rectangles, and read/write accesses 
  with plain edges. 
  For visualization purposes, the process tree is represented too, with
  dashed edges. Every node in the graph is labeled using (1) a process id
  created by our reconstruction, (2) the name of the executable run by the
  process or the file name. Process 0 is the initial call to the
  \texttt{sh} interpreter; the other processes are the calls to FSL bet,
  stats and maths made in Listing~\ref{listing:sample-script}. Process 73
  is forked by process 2: it was captured by \reprozip while it did not
  appear in Listing~\ref{listing:sample-script}.
}
  \label{fig:simple_script}
\end{figure}

@@

We capture temporary files by replacing every
process P by a wrapper that first calls P and then saves the produced
temporary files to a read-only directory. This process replacement is done by pre-pending
 to the \texttt{PATH} environment
variable a directory that contains a wrapper script named after the executable
called by P.

Files written by multiple processes are disambiguated using a similar technique. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, we first check that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, we
 replace every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way,
 successive versions of F are preserved for comparison. We finally
 update the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1 (Figure~\ref{fig:provenance-dag}). This operation also makes the provenance graph
 acyclic, since we assumed that a process could only release a single version of a file.

@@

After capturing transient files, we re-run the pipeline
step by step in Condition 2 to label processes. The output files
produced by a process in Condition 2 are compared to the ones produced
in Condition 1. If no differences are found, the process is marked as
reproducible. Otherwise, the process is marked as non-reproducible and the
output files produced in Condition 1 are copied to Condition 2, to ensure
that differences do not propagate further in the pipeline. Processes are
instrumented transparently, through a modification of the \texttt{PATH}
variable similar to the one described previously. By default, differences
in output files are identified by comparing file checksums. Other
comparison functions can also be defined for specific file types, for
instance to ignore file headers or file sections containing timestamps.
\toolname finally creates a labeled
provenance graph highlighting non-reproducible processes.

Figure~\ref{fig:iterations} illustrates the incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. 
At step 1, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
as it produces files with differences. Subsequently, the files produced by \texttt{73\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
Condition 1.
At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
as reproducible (green) as it produced files without differences.
At step 3, \texttt{76\#fslstats} is also labeled as 
reproducible as it didn't produce any file.

@@

@@

Propagation is very important.
